State of the Practice for Lattice Boltzmann Method Software

### 2022-04-21 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revisions to abstract, first two paragraphs of intro

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP). Our ranking was roughly
    consistent with ranking by GitHub stars, suggesting at least a partial
    correlation between the use of best practices and popularity. We find the
    state of the practice to be healthy, with 67\% of the measured packages
    ranking in the top five for at least one quality, the majority of LBM
    generated artifacts corresponding to general recommendations from research
    software developers, common use of version control (67\% of
    packages) and adoption of a quasi-agile development process.  Areas of best
    practice to potentially improve include adoption of continuous integration,
    API documentation and enforcement of programming style guides.  We
    interviewed four developers to gain insight into their current pain points.
    Identified challenges include lack of development time, lack of funding, and
    difficulty with ensuring correctness.  Developers are addressing these pain
    points by designing for change, circumventing the oracle problem and
    prioritizing documentation and usability. For future improvements we suggest
    the following: employing linters, conducting rigorous peer reviews, writing
    and submitting more papers on software, growing the number of contributors
    by following current recommendations for open source projects, and
    augmenting the theory manuals to include more requirements relevant
    information.

@@

We analyze the development of Computational Fluid Dynamics (CFD) software
packages that use the Lattice Boltzmann Method (LBM). LBM packages form a family
of algorithms for simulating single-phase and multiphase fluid flows, often
incorporating additional physical complexities \citep{chen1998lattice}, such as
reflective and non-reflective boundaries. LBM considers the behaviour of a
collection of particles as a single unit at the mesoscopic scale, which lies
between the nanoscopic and microscopic scales. LBM solvers predict the
positional probability of a collection of particles moving through a lattice
structure following a two step process: i) streaming, where the particles move
along the lattice via links; and, ii) colliding, where energy and momentum is
transferred among particles that collide \citep{bao2011lattice}. As an example
of an important application of LBM, Figure~\ref{Fig_coarctation} shows the flow
in a defective aorta pre and post medical intervention. LBM has several
advantages over conventional CFD methods, including a simple calculation
procedure, improved parallelization, and robust handling of complex geometries
\citep{ganji2015application}.

@@

A small sample of important applications of LBM include the following: designing
fuel cells \citep{ZhangEtAl2018}, modelling groundwater flow
\citep{AnwarAndSukop2009}, and analyzing the flow of blood in the cardiovascular
system \citep{SadeghiEtAl2022b, SadeghiEtAl2022, SadeghiEtAl2020}.  As these
examples illustrate, LBM can be used for tackling problems that impact such
areas as environmental policy, manufacturing, and health and safety.  Given the
important applications of LBM, users of the libraries will be concerned with
software qualities like reliability, robustness, reproducibility, performance,
correctness and verifiability.  Since their time is valuable, developers that
create or modify LBM libraries will have additional quality concerns, including
maintainability, reusability, and understandability.  With the quality
considerations in mind, and the potentially overwhelming number of choices for
existing LBM libraries, the time is right to assess the state of the practice.
The goal of this report is to analyze the current state of LBM software
development and compare it to the development practices employed for research
software in general.  We want to highlight success stories that LBM developers
can share amongst their community, and with the broader research software
community, while at the same time looking for areas for potential future
improvement.

>>>>

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of 20 questions). After measuring we ranked the software using
    the Analytic Hierarchy Process (AHP). Our ranking was roughly consistent
    with GitHub stars ranking, suggesting at least a partial correlation
    between the use of best practices and popularity. We find the state of the
    practice to be healthy, with 67\% of the measured packages ranking in the
    top five for at least one quality, the majority of LBM generated artifacts
    corresponding to general recommendations from research software developers,
    common use of version control (67\% of packages) and the adoption of a
    quasi-agile development process.  Areas of best practice to potentially
    improve include adoption of continuous integration, API documentation and
    enforcement of programming style guides.  We interviewed four developers to
    gain insight into their current pain points. Identified challenges include
    lack of development time, lack of funding, and difficulty with ensuring
    correctness.  Developers are addressing these pain points by designing for
    change, circumventing the oracle problem and prioritizing documentation and
    usability. For future improvements we suggest the following: employing
    linters, conducting rigorous peer reviews, writing and submitting more
    papers on software, growing the number of contributors by following current
    recommendations for open source projects, and augmenting the theory manuals
    to include more requirements relevant information.

@@

We analyze the development of Computational Fluid Dynamics (CFD) software
packages that use the Lattice Boltzmann Method (LBM). LBM packages form a family
of algorithms for simulating single-phase and multiphase fluid flows, often
incorporating additional physical complexities \citep{chen1998lattice}, such as
reflective and non-reflective boundaries. LBM considers the behaviour of a
collection of particles as a single unit at the mesoscopic scale, which lies
between the nanoscopic and microscopic scales. LBM solvers predict the
positional probability of a collection of particles moving through a lattice
structure following a two-step process: i) streaming, where the particles move
along the lattice via links; and, ii) colliding, where energy and momentum are
transferred among particles that collide \citep{bao2011lattice}. As an example
of an important application of LBM, Figure~\ref{Fig_coarctation} shows the flow
in a defective aorta pre and post-medical intervention. LBM has several
advantages over conventional CFD methods, including a simple calculation
procedure, improved parallelization, and robust handling of complex geometries
\citep{ganji2015application}.

@@

A small sample of important applications of LBM includes the following:
designing fuel cells \citep{ZhangEtAl2018}, modelling groundwater flow
\citep{AnwarAndSukop2009}, and analyzing the flow of blood in the cardiovascular
system \citep{SadeghiEtAl2022b, SadeghiEtAl2022, SadeghiEtAl2020}.  As these
examples illustrate, LBM can be used for tackling problems that impact such
areas as environmental policy, manufacturing, and health and safety.  Given the
important applications of LBM, users of the libraries will be concerned with
software qualities like reliability, robustness, reproducibility, performance,
correctness and verifiability.  Since their time is valuable, developers that
create or modify LBM libraries will have additional quality concerns, including
maintainability, reusability, and understandability.  With the quality
considerations in mind, and the potentially overwhelming number of choices for
existing LBM libraries, the time is right to assess the state of the practice.
The goal of this report is to analyze the current state of LBM software
development and compare it to the development practices employed for research
software in general.  We want to highlight success stories that LBM developers
can share amongst their community, and with the broader research software
community, while at the same time looking for areas for potential future
improvement.
